---
title: "imcExperiment: containerization of IMC data"
author: "Anthony Colombo, Erik Gerdtsson"
output: 
  html_vignette:
  toc: true
  number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{arkasData}
  %\VignetteIndexEntry{Arkas: Repetitive Elements Quantification In Much Less Time}
  \usepackage[utf8]{inputenc}   
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# example
 This shows how to construct and IMCexperiment object.
  need to make sure that the number of rows from spatial match number of rows of expression.
  the sizes need to be checked.

```{r}

  x<-imcExperiment(cellIntensity=matrix(1,nrow=10,ncol=10),
	spatial=matrix(1,nrow=10,ncol=2),
	neighborHood=matrix(1,nrow=10,ncol=10),
	network=matrix(1,nrow=10,ncol=10),
	uniqueLabel=rep("A",10),
	ROIID=data.frame(ROIID=rep("A",10)))

 x<-imcExperiment(cellIntensity=matrix(1,nrow=7,ncol=10),
	spatial=matrix(1,nrow=10,ncol=2),
	neighborHood=matrix(1,nrow=10,ncol=10),
	network=matrix(1,nrow=10,ncol=10),
	uniqueLabel=rep("A",10),
	ROIID=data.frame(ROIID=rep("A",10)))


 getSpatial(x)<-matrix(0,10,2)
 cellIntensity(x)

  cellIntensity(x)<-asinh(cellIntensity(x))
  cellIntensity(x)


``` 

### From histoCAT to R
* The IMC container contains cellIntensity, spatial coordinates, neighborhoods, networking assignments, unique cell labeling, and ROIID.
* each slot has an accessor, and the ROIID is used to set the rowData of your object.
```{r}
  ### load the data from package.
  library(imcExperiment)
 ##load the data 1000 cells from IMC experiment.
  data(data)
  dim(data)
  ##output from histoCAT to R
  expr<-data[,3:36]
  normExp<-percentilenormalize(data=expr,percentile=0.99)
  normExp<-as.matrix(normExp)
  boxplot(normExp)

  ##spatial component
  spatial<-(data[,c("X_position","Y_position")])
  spatial<-as.matrix(spatial)
 ##uniqueLabel
  uniqueLabel<-paste0(data[,"ImageId"],"_",data[,"CellId"])

 phenotypes<-matrix(data[,"Phenograph4732517031"])

   x<-imcExperiment(cellIntensity=normExp,
	spatial=spatial,
	neighborHood=phenotypes,
	network=phenotypes,
	uniqueLabel=uniqueLabel,
	ROIID=NULL)
  ### create  phenotypes via Rphenograph
  ##run phenograph
  require(Rphenograph)
  phenos<-Rphenograph(cellIntensity(x),k=35)
   pheno.labels<-as.numeric(membership(phenos[[2]]))
   getNetwork(x)<-matrix(pheno.labels)
  head(getNetwork(x))
  ##plot phenograph
  plot_clustering_heatmap_wrapper(myExperiment=x)

 ###testing each accessor.
  ## need to get each slot.
  ##cell intensity
   head(cellIntensity(x))
   cellIntensity(x)<-cellIntensity(x)
   all(cellIntensity(x)==normExp)
    all(assays(x)$exprs==cellIntensity(x))
   
   ##spatial coordinates
   head(getSpatial(x))
    getSpatial(x)<-spatial
     all(getSpatial(x)==spatial)
   ##
     
     ##neighborhood cell assignments.
      head(getNetwork(x))
      all(getNetwork(x)==as.matrix(pheno.labels))
      
      ##uniqueLabels
      head(getLabel(x))
      all(getLabel(x)==uniqueLabel)
      
      

      

```  

### Read in histoCat data into R
* read in histoCat data for customized analysis.

```{r}
### reading in a directory of histoCAT csv files.

 ##need to reconstruct the fold revised IMC data.
  ##merges a folder full of histoCAT csv files.
  dataSet<-NULL
 for(i in c("case8.csv","case5.csv")){
  message(i)
    fpath <- system.file("extdata", i, package="imcExperiment")
    case1<-read.csv(fpath)
   proteins<-colnames(case1)[grepl("Cell_",colnames(case1))]
    case1<-case1[,c("ImageId","CellId","X_position","Y_position",proteins)]
    dataSet<-rbind(dataSet,case1)
  }
 
 expr<-dataSet[,proteins]
  normExp<-percentilenormalize(data=expr,percentile=0.99)
  normExp<-as.matrix(normExp)
  boxplot(normExp)

  ##spatial component
  spatial<-(dataSet[,c("X_position","Y_position")])
  spatial<-as.matrix(spatial)
 ##uniqueLabel
  uniqueLabel<-paste0(dataSet[,"ImageId"],"_",dataSet[,"CellId"])

  ##not yet assigned
  phenotypes<-matrix(0,nrow(dataSet),1)
  ROIID=data.frame(ROIID=dataSet[,"ImageId"])
   x<-imcExperiment(cellIntensity=normExp,
        spatial=spatial,
        neighborHood=phenotypes,
        network=phenotypes,
        uniqueLabel=uniqueLabel,
	ROIID=ROIID)
  ##phenotype
  ##run phenograph
  require(Rphenograph)
  phenos<-Rphenograph(cellIntensity(x),k=35)
   pheno.labels<-as.numeric(membership(phenos[[2]]))
   getNetwork(x)<-matrix(pheno.labels)
   head( getNetwork(x))
  ##plot phenograph
  plot_clustering_heatmap_wrapper(myExperiment=x)


 ##store the ROIID in the metadata columns.
  ##access the unique cell labels.
 head( getLabel(x))
  
  ##you can append more clinical features to the columns of the sampleDat data.frame.
  H<-imcExperimentToHyperFrame(imcExperiment=x)
  helper<-function(pp=NULL,i=NULL,j=NULL){
  ps<-split(pp)
  nnd<-nncross(ps[[i]],ps[[j]])
  }
 ## 1-NN analysis.
  ## compare cluster 10 to cluster 9
   eachNND<-with(H,helper(pp=point,i="10",j="8"))
 ## first choose 2 clusters to compare.
  sapply(eachNND,function(x) mean(x[,"dist"]))




```
